\bf The forestecology R package for fitting and assessing models of interspecies competitive effects on the growth of trees

### 2021-03-04 paper/forestecology.Rmd
more formatting

authors: 
- name: Albert Y. Kim 
  affiliation: Program in Statistical & Data Sciences, Smith College

@@

keywords:
- forest ecology, competition, R, Rstats, tidyverse, sf, cross-validation, spatial statistics

spacing: double

@@

# knitr settings
opts_chunk$set(
  # Code output:
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  cache = TRUE,
  # Figure:
  out.width = "100%",
  fig.path = "Figures/",
  fig.width = 16 / 2.5,
  fig.height = 9 / 2.5,
  fig.align = "center",
  fig.show = "hold",
  # Etc:
  collapse = TRUE,
  comment = "##"
)

@@

# Set output width
options(
  width = 70, 
  # Decimal precision
  digits = 3, 
  # Number of rows for tibble printing
  tibble.print_max = 5, tibble.print_min = 5
)
```

@@

## Step 1: Compute the growth of trees based on census data

@@

We load both 2008 and 2014 SCBI census data `.csv` files as they existed on GitHub on November 20, 2020. After selecting only the relevant variables, we perform a few additional data wrangling steps: convert the character variable with the date of measurement to be of explicit type `date`, convert DBH to be in centimeters^[A rule of thumb to determine the units of DBH is to check if the smallest non-zero and non-missing measurement is 1 or 10. If the former, then centimeters. If the later, then millimeters. This is because ForestGEO protocols state that only trees with DBH greater or equal to 1cm should be included in censuses.] Furthermore, in order to speed up computation for purposes of this example, we only consider a 9 ha subsection of the 25.6 ha of the SCBI site: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.

@@

The second element of spatial information are blocks corresponding to folds of a spatial cross-validation algorithm used to estimate out-of-sample model error. Conventional cross-validation algorithms assign observations to folds by randomly resampling individual observations. However, many of these algorithms assume that the observations are independent of each other. In the case of forest census data, observations exhibit spatial autocorrelation. We therefore incorporate this spatial dependence into the cross-validation algorithm with our spatial blocks of trees [@roberts_cross-validation_2017 @pohjankukka_estimating_2017]
In the example below, we first manually define four folds that partition the study region as an `sf` object. We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct fold (saved in the `foldID` variable) [@valavi_blockcv_2019]. ^[In the Appendix we present an example where the folds themselves are also created using the `spatialBlock()` function given a specified `cv_block_size`.]

@@

The next step in our analysis sequence is to identify all focal trees and their corresponding competitor trees. More specifically, it identifies all trees that are not part of the buffer region, have a valid `growth` measurement, and have at least one neighbor within `r comp_dist` of it.  The `create_focal_vs_comp()` functions performs these tasks and returns a new data frame of type `sf`. On top of the previous arguments `comp_dist` defining the competition neighborhood and `id` indicating which variable uniquely identifies each tree-stem, this function also requires an `sf` object representation of the spatial cross-validation blocks/folds. In this example, the blocks were manually encoded in `blocks_scbi` by specifying it's vertices in Section \ref{spatial-information}^[We present an alternative method for defining spatial cross-validation blocks is using the `spatialBlock()` function from the `blockCV` package in the Appendix]. We present the resulting data frame below with the `foldID` variable omitted for compactness of presentation.

@@

The resulting RMSE of `r model_rmse_shuffle` based on the permutation test is `r ifelse(model_rmse_shuffle < model_rmse, "smaller", "larger")` than the earlier RMSE of `r model_rmse`, suggesting that models that do incorporate competitor species identity better fit the data. We conduct an analysis of the full SCBI plot in Section below.

@@

The resulting RMSE of `r model_rmse_cv` computed using cross-validation is `r ifelse(model_rmse_cv > model_rmse, "larger", "smaller")` than the earlier RMSE of `r model_rmse`, suggesting that models that do not take the inherent spatial autocorrelation of the data into account generate error estimates that are overly optimistic; in our case RMSE's that are too low.

@@

We intend to archive all data and code on GitHub at the following repository https://github.com/rudeboybert/forestecology. This repository will be versioned and archived on Zenodo upon acceptance.

>>>>

authors: 
- name: Albert Y. Kim 
  affiliation: Program in Statistical & Data Sciences, Smith College
  thanks: "Assistant Professor, Statistical & Data Sciences, Smith College, Northampton, MA 01063 (e-mail: akim04@smith.edu)."

@@

keywords:
- forest ecology, inter-species competition, tree growth, R, Rstats, tidyverse, sf, cross-validation, spatial statistics

@@

# knitr settings
opts_chunk$set(
  # Code output:
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  cache = TRUE,
  # Figure:
  out.width = "100%",
  fig.path = "Figures/",
  fig.width = 16 / 2.5,
  fig.height = 9 / 2.5,
  fig.align = "center",
  fig.show = "hold",
  # Etc:
  collapse = TRUE,
  comment = "##",
  tidy.opts = list(width.cutoff=72),
  tidy = TRUE
)

@@

# Set output width
options(
  width = 72, 
  # Decimal precision
  digits = 3, 
  # Number of rows for tibble printing
  tibble.print_max = 5, 
  tibble.print_min = 5
)
```

@@

## Step 1: Compute the growth of trees based on census data {compute-growth}

@@

We load both 2008 and 2014 SCBI census data `.csv` files as they existed on GitHub on November 20, 2020 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020]. After selecting only the relevant variables, we perform a few additional data wrangling steps: convert the character variable with the date of measurement to be of explicit type `date`, convert DBH to be in centimeters^[A rule of thumb to determine the units of DBH is to check if the smallest non-zero and non-missing measurement is 1 or 10. If the former, then centimeters. If the later, then millimeters. This is because ForestGEO protocols state that only trees with DBH greater or equal to 1cm should be included in censuses.] Furthermore, in order to speed up computation for purposes of this example, we only consider a 9 ha subsection of the 25.6 ha of the SCBI site: `gx` from 0--300 instead of 0--400 and `gy` from 300--600 instead of 0--640.

@@

The second element of spatial information are blocks corresponding to folds of a spatial cross-validation algorithm used to estimate out-of-sample model error. Conventional cross-validation algorithms assign observations to folds by randomly resampling individual observations. However, many of these algorithms assume that the observations are independent of each other. In the case of forest census data, observations exhibit spatial autocorrelation. We therefore incorporate this spatial dependence into the cross-validation algorithm with our spatial blocks of trees [@roberts_cross-validation_2017 @pohjankukka_estimating_2017]
In the example below, we first manually define four folds that partition the study region as an `sf` object. We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct fold (saved in the `foldID` variable) [@valavi_blockcv_2019]. ^[In the Supporting Information we present an example where the folds themselves are also created using the `spatialBlock()` function given a specified `cv_block_size`.]

@@

The next step in our analysis sequence is to identify all focal trees and their corresponding competitor trees. More specifically, it identifies all trees that are not part of the buffer region, have a valid `growth` measurement, and have at least one neighbor within `r comp_dist` of it.  The `create_focal_vs_comp()` functions performs these tasks and returns a new data frame of type `sf`. On top of the previous arguments `comp_dist` defining the competition neighborhood and `id` indicating which variable uniquely identifies each tree-stem, this function also requires an `sf` object representation of the spatial cross-validation blocks/folds. In this example, the blocks were manually encoded in `blocks_scbi` by specifying it's vertices in Section \ref{spatial-information}^[We present an alternative method for defining spatial cross-validation blocks is using the `spatialBlock()` function from the `blockCV` package in the Supporting Information.]. We present the resulting data frame below with the `foldID` variable omitted for compactness of presentation.

@@

The resulting RMSE of `r model_rmse_shuffle` based on the permutation test is `r ifelse(model_rmse_shuffle < model_rmse, "smaller", "larger")` than the earlier RMSE of `r model_rmse`, suggesting that models that do incorporate competitor species identity better fit the data. We conduct an analysis of the full SCBI plot in Section \ref{results} below.

@@

The resulting RMSE of `r model_rmse_cv` computed using cross-validation is `r ifelse(model_rmse_cv > model_rmse, "larger", "smaller")` than the earlier RMSE of `r model_rmse`, suggesting that models that do not take the inherent spatial autocorrelation of the data into account generate error estimates that are overly optimistic; in our case RMSE's that are too low. We conduct an analysis of the full SCBI plot in Section \ref{results} below.

# Results {#results}

Figure \ref{fig:scbi-simulation}

```{r scbi-simulation, out.width="100%", fig.cap="Root mean squared error of models for standard, permuted, and spatial cross-validated error estimates.", echo = FALSE}
include_graphics("simulation_results/2021-03-03_scbi_49_shuffles.png")
```

@@

We intend to archive all data and source code for the `forestecology` package as well as this manuscript on GitHub at https://github.com/rudeboybert/forestecology. This repository will be versioned and archived on Zenodo upon acceptance. The 2008 and 2014 Smithsonian Conservation Biology Institute census data loaded in Section \ref{compute-growth} and saved in `scbi.stem2.csv` and `scbi.stem3.csv` are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files and have been versioned and archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].

### 2021-03-04 paper/forestecology.tex
more formatting

  \author{
        Albert Y. Kim \\
    Program in Statistical \& Data Sciences, Smith College\\
     and \\     David N. Allen \\
    Biology Department, Middlebury College\\
     and \\     Simon P. Couch \\
    Mathematics Department, Reed College\\
      }
  \maketitle
} \fi

@@

\noindent%
{\it Keywords:} forest ecology, competition, R, Rstats, tidyverse, sf, cross-validation,
spatial statistics
\vfill

@@

\hypertarget{step-1-compute-the-growth-of-trees-based-on-census-data}{%
\subsection{Step 1: Compute the growth of trees based on census
data}\label{step-1-compute-the-growth-of-trees-based-on-census-data}}

@@

We load both 2008 and 2014 SCBI census data \texttt{.csv} files as they
existed on GitHub on November 20, 2020. After selecting only the
relevant variables, we perform a few additional data wrangling steps:
convert the character variable with the date of measurement to be of
explicit type \texttt{date}, convert DBH to be in centimeters\footnote{A
  rule of thumb to determine the units of DBH is to check if the
  smallest non-zero and non-missing measurement is 1 or 10. If the
  former, then centimeters. If the later, then millimeters. This is
  because ForestGEO protocols state that only trees with DBH greater or
  equal to 1cm should be included in censuses.} Furthermore, in order to
speed up computation for purposes of this example, we only consider a 9
ha subsection of the 25.6 ha of the SCBI site: \texttt{gx} from 0--300
instead of 0--400 and \texttt{gy} from 300--600 instead of 0--640.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{growth_scbi <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
    \DataTypeTok{id =} \StringTok{"stemID"}
\NormalTok{  )}
\NormalTok{growth_scbi}
\CommentTok{## Simple feature collection with 7954 features and 8 fields}
\CommentTok{## geometry type:  POINT}
\CommentTok{## dimension:      XY}
\CommentTok{## bbox:           xmin: 0.2 ymin: 300 xmax: 300 ymax: 600}
\CommentTok{## CRS:            NA}
\CommentTok{## # A tibble: 7,954 x 9}
\CommentTok{##   stemID sp     dbh1 codes1 status  dbh2 codes2 growth   geometry}
\CommentTok{##    <dbl> <fct> <dbl> <chr>  <chr>  <dbl> <chr>   <dbl>    <POINT>}
\CommentTok{## 1      4 nysy  13.6  M      A       14.2 M       0.103 (14.2 428)}
\CommentTok{## 2      5 havi   8.8  M      A        9.6 M;P     0.150  (9.4 436)}
\CommentTok{## 3      6 havi   3.25 NULL   A        4   M       0.140  (1.3 434)}
\CommentTok{## 4     77 qual  65.2  M      A       66   M       0.141 (34.7 307)}
\CommentTok{## 5     79 tiam  47.7  M      A       46.8 M      -0.161   (40 381)}
\CommentTok{## # ... with 7,949 more rows}
\end{Highlighting}
\end{Shaded}

@@

The second element of spatial information are blocks corresponding to
folds of a spatial cross-validation algorithm used to estimate
out-of-sample model error. Conventional cross-validation algorithms
assign observations to folds by randomly resampling individual
observations. However, many of these algorithms assume that the
observations are independent of each other. In the case of forest census
data, observations exhibit spatial autocorrelation. We therefore
incorporate this spatial dependence into the cross-validation algorithm
with our spatial blocks of trees \citep[
\citet{pohjankukka_estimating_2017}]{roberts_cross-validation_2017} In
the example below, we first manually define four folds that partition
the study region as an \texttt{sf} object. We then use the output of the
\texttt{spatialBlock()} function from the \texttt{blockCV} package to
associate each tree in \texttt{growth\_scbi} to the correct fold (saved
in the \texttt{foldID} variable) \citep{valavi_blockcv_2019}. \footnote{In
  the Appendix we present an example where the folds themselves are also
  created using the \texttt{spatialBlock()} function given a specified
  \texttt{cv\_block\_size}.}

@@

The next step in our analysis sequence is to identify all focal trees
and their corresponding competitor trees. More specifically, it
identifies all trees that are not part of the buffer region, have a
valid \texttt{growth} measurement, and have at least one neighbor within
7.5 of it. The \texttt{create\_focal\_vs\_comp()} functions performs
these tasks and returns a new data frame of type \texttt{sf}. On top of
the previous arguments \texttt{comp\_dist} defining the competition
neighborhood and \texttt{id} indicating which variable uniquely
identifies each tree-stem, this function also requires an \texttt{sf}
object representation of the spatial cross-validation blocks/folds. In
this example, the blocks were manually encoded in \texttt{blocks\_scbi}
by specifying it's vertices in Section
\ref{spatial-information}\footnote{We present an alternative method for
  defining spatial cross-validation blocks is using the
  \texttt{spatialBlock()} function from the \texttt{blockCV} package in
  the Appendix}. We present the resulting data frame below with the
\texttt{foldID} variable omitted for compactness of presentation.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{comp_bayes_lm_scbi}
\CommentTok{## Bayesian linear regression model parameters with a multivariate Normal likelihood. See ?comp_bayes_lm for details:}
\CommentTok{## }
\CommentTok{##   parameter_type           prior posterior}
\CommentTok{## 1 Inverse-Gamma on sigma^2 a_0   a_star   }
\CommentTok{## 2 Inverse-Gamma on sigma^2 b_0   b_star   }
\CommentTok{## 3 Multivariate t on beta   mu_0  mu_star  }
\CommentTok{## 4 Multivariate t on beta   V_0   V_star   }
\CommentTok{## }
\CommentTok{## Model formula:}
\CommentTok{## growth ~ sp + dbh + dbh * sp + acne * sp + acpl * sp + acru * sp + acsp * sp + aial * sp + amar * sp + astr * sp + beth * sp + caca * sp + caco * sp + cade * sp + cagl * sp + caovl * sp + casp * sp + cato * sp + ceca * sp + ceoc * sp + chvi * sp + coal * sp + coam * sp + cofl * sp + crpr * sp + crsp * sp + divi * sp + elum * sp + eual * sp + fagr * sp + fram * sp + frni * sp + frpe * sp + frsp * sp + havi * sp + ilve * sp + juci * sp + juni * sp + juvi * sp + libe * sp + litu * sp + loma * sp + nysy * sp + pato * sp + pipu * sp + pist * sp + pivi * sp + ploc * sp + prav * sp + prpe * sp + prse * sp + prsp * sp + qual * sp + quco * sp + qufa * sp + qumi * sp + qupr * sp + quru * sp + qusp * sp + quve * sp + rhpe * sp + romu * sp + rops * sp + rual * sp + rupe * sp + ruph * sp + saal * sp + saca * sp + tiam * sp + ulam * sp + ulru * sp + ulsp * sp + unk * sp + viac * sp + vipr * sp + vire * sp}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi}
\CommentTok{## # A tibble: 6,296 x 8}
\CommentTok{##   focal_ID focal_sp   dbh foldID   geometry growth}
\CommentTok{##      <dbl> <fct>    <dbl> <fct>     <POINT>  <dbl>}
\CommentTok{## 1        4 nysy     13.6  1      (14.2 428)  0.103}
\CommentTok{## 2        5 havi      8.8  1       (9.4 436)  0.150}
\CommentTok{## 3       79 tiam     47.7  1        (40 381) -0.161}
\CommentTok{## 4       80 caca      5.15 1      (38.7 422)  0.253}
\CommentTok{## 5       96 libe      2.3  1        (60 310)  0.262}
\CommentTok{## # ... with 6,291 more rows, and 2 more variables: comp <list>,}
\CommentTok{## #   growth_hat <dbl>}
\end{Highlighting}
\end{Shaded}

@@

The resulting RMSE of 0.131 based on the permutation test is larger than
the earlier RMSE of 0.128, suggesting that models that do incorporate
competitor species identity better fit the data. We conduct an analysis
of the full SCBI plot in Section below.

@@

The resulting RMSE of 0.14 computed using cross-validation is larger
than the earlier RMSE of 0.128, suggesting that models that do not take
the inherent spatial autocorrelation of the data into account generate
error estimates that are overly optimistic; in our case RMSE's that are
too low.

@@

We intend to archive all data and code on GitHub at the following
repository \url{https://github.com/rudeboybert/forestecology}. This
repository will be versioned and archived on Zenodo upon acceptance.

>>>>

  \author{
        Albert Y. Kim \thanks{Assistant Professor, Statistical \& Data Sciences, Smith College,
Northampton, MA 01063 (e-mail:
\href{mailto:akim04@smith.edu}{\nolinkurl{akim04@smith.edu}}).} \\
    Program in Statistical \& Data Sciences, Smith College\\
     and \\     David N. Allen \\
    Biology Department, Middlebury College\\
     and \\     Simon P. Couch \\
    Mathematics Department, Reed College\\
      }
  \maketitle
} \fi

@@

\noindent%
{\it Keywords:} forest ecology, inter-species competition, tree growth, R, Rstats,
tidyverse, sf, cross-validation, spatial statistics
\vfill

@@

\hypertarget{step-1-compute-the-growth-of-trees-based-on-census-data-compute-growth}{%
\subsection{Step 1: Compute the growth of trees based on census data
\{compute-growth\}}\label{step-1-compute-the-growth-of-trees-based-on-census-data-compute-growth}}

@@

We load both 2008 and 2014 SCBI census data \texttt{.csv} files as they
existed on GitHub on November 20, 2020
\citep{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}. After
selecting only the relevant variables, we perform a few additional data
wrangling steps: convert the character variable with the date of
measurement to be of explicit type \texttt{date}, convert DBH to be in
centimeters\footnote{A rule of thumb to determine the units of DBH is to
  check if the smallest non-zero and non-missing measurement is 1 or 10.
  If the former, then centimeters. If the later, then millimeters. This
  is because ForestGEO protocols state that only trees with DBH greater
  or equal to 1cm should be included in censuses.} Furthermore, in order
to speed up computation for purposes of this example, we only consider a
9 ha subsection of the 25.6 ha of the SCBI site: \texttt{gx} from 0--300
instead of 0--400 and \texttt{gy} from 300--600 instead of 0--640.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{growth_scbi <-}
\StringTok{  }\KeywordTok{compute_growth}\NormalTok{(}
    \DataTypeTok{census_1 =}\NormalTok{ census_}\DecValTok{2013}\NormalTok{_scbi,}
    \DataTypeTok{census_2 =}\NormalTok{ census_}\DecValTok{2018}\NormalTok{_scbi }\OperatorTok{%>%}\StringTok{ }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{str_detect}\NormalTok{(codes, }\StringTok{"R"}\NormalTok{)),}
    \DataTypeTok{id =} \StringTok{"stemID"}
\NormalTok{  )}
\NormalTok{growth_scbi}
\CommentTok{## Simple feature collection with 7954 features and 8 fields}
\CommentTok{## geometry type:  POINT}
\CommentTok{## dimension:      XY}
\CommentTok{## bbox:           xmin: 0.2 ymin: 300 xmax: 300 ymax: 600}
\CommentTok{## CRS:            NA}
\CommentTok{## # A tibble: 7,954 x 9}
\CommentTok{##   stemID sp     dbh1 codes1 status  dbh2 codes2 growth}
\CommentTok{##    <dbl> <fct> <dbl> <chr>  <chr>  <dbl> <chr>   <dbl>}
\CommentTok{## 1      4 nysy  13.6  M      A       14.2 M       0.103}
\CommentTok{## 2      5 havi   8.8  M      A        9.6 M;P     0.150}
\CommentTok{## 3      6 havi   3.25 NULL   A        4   M       0.140}
\CommentTok{## 4     77 qual  65.2  M      A       66   M       0.141}
\CommentTok{## 5     79 tiam  47.7  M      A       46.8 M      -0.161}
\CommentTok{## # ... with 7,949 more rows, and 1 more variable: geometry <POINT>}
\end{Highlighting}
\end{Shaded}

@@

The second element of spatial information are blocks corresponding to
folds of a spatial cross-validation algorithm used to estimate
out-of-sample model error. Conventional cross-validation algorithms
assign observations to folds by randomly resampling individual
observations. However, many of these algorithms assume that the
observations are independent of each other. In the case of forest census
data, observations exhibit spatial autocorrelation. We therefore
incorporate this spatial dependence into the cross-validation algorithm
with our spatial blocks of trees \citep[
\citet{pohjankukka_estimating_2017}]{roberts_cross-validation_2017} In
the example below, we first manually define four folds that partition
the study region as an \texttt{sf} object. We then use the output of the
\texttt{spatialBlock()} function from the \texttt{blockCV} package to
associate each tree in \texttt{growth\_scbi} to the correct fold (saved
in the \texttt{foldID} variable) \citep{valavi_blockcv_2019}. \footnote{In
  the Supporting Information we present an example where the folds
  themselves are also created using the \texttt{spatialBlock()} function
  given a specified \texttt{cv\_block\_size}.}

@@

The next step in our analysis sequence is to identify all focal trees
and their corresponding competitor trees. More specifically, it
identifies all trees that are not part of the buffer region, have a
valid \texttt{growth} measurement, and have at least one neighbor within
7.5 of it. The \texttt{create\_focal\_vs\_comp()} functions performs
these tasks and returns a new data frame of type \texttt{sf}. On top of
the previous arguments \texttt{comp\_dist} defining the competition
neighborhood and \texttt{id} indicating which variable uniquely
identifies each tree-stem, this function also requires an \texttt{sf}
object representation of the spatial cross-validation blocks/folds. In
this example, the blocks were manually encoded in \texttt{blocks\_scbi}
by specifying it's vertices in Section
\ref{spatial-information}\footnote{We present an alternative method for
  defining spatial cross-validation blocks is using the
  \texttt{spatialBlock()} function from the \texttt{blockCV} package in
  the Supporting Information.}. We present the resulting data frame
below with the \texttt{foldID} variable omitted for compactness of
presentation.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{comp_bayes_lm_scbi}
\CommentTok{## Bayesian linear regression model parameters with a multivariate Normal likelihood. See ?comp_bayes_lm for details:}
\CommentTok{## }
\CommentTok{##   parameter_type           prior posterior}
\CommentTok{## 1 Inverse-Gamma on sigma^2 a_0   a_star   }
\CommentTok{## 2 Inverse-Gamma on sigma^2 b_0   b_star   }
\CommentTok{## 3 Multivariate t on beta   mu_0  mu_star  }
\CommentTok{## 4 Multivariate t on beta   V_0   V_star   }
\CommentTok{## }
\CommentTok{## Model formula:}
\CommentTok{## growth ~ sp + dbh + dbh * sp + acne * sp + acru * sp + amar * sp + astr * sp + caca * sp + caco * sp + cade * sp + cagl * sp + caovl * sp + cato * sp + ceca * sp + ceoc * sp + chvi * sp + cofl * sp + crpr * sp + crsp * sp + divi * sp + elum * sp + fagr * sp + fram * sp + frni * sp + frpe * sp + havi * sp + ilve * sp + juci * sp + juni * sp + libe * sp + litu * sp + nysy * sp + pist * sp + pivi * sp + ploc * sp + prav * sp + prse * sp + qual * sp + quco * sp + qufa * sp + qumi * sp + qupr * sp + quru * sp + quve * sp + rops * sp + saal * sp + saca * sp + tiam * sp + ulam * sp + ulru * sp + unk * sp + vipr * sp}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_scbi}
\CommentTok{## # A tibble: 6,296 x 8}
\CommentTok{##   focal_ID focal_sp   dbh foldID   geometry growth comp             }
\CommentTok{##      <dbl> <fct>    <dbl> <fct>     <POINT>  <dbl> <list>           }
\CommentTok{## 1        4 nysy     13.6  1      (14.2 428)  0.103 <tibble [20 x 4]>}
\CommentTok{## 2        5 havi      8.8  1       (9.4 436)  0.150 <tibble [32 x 4]>}
\CommentTok{## 3       79 tiam     47.7  1        (40 381) -0.161 <tibble [20 x 4]>}
\CommentTok{## 4       80 caca      5.15 1      (38.7 422)  0.253 <tibble [12 x 4]>}
\CommentTok{## 5       96 libe      2.3  1        (60 310)  0.262 <tibble [14 x 4]>}
\CommentTok{## # ... with 6,291 more rows, and 1 more variable: growth_hat <dbl>}
\end{Highlighting}
\end{Shaded}

@@

The resulting RMSE of 0.131 based on the permutation test is larger than
the earlier RMSE of 0.128, suggesting that models that do incorporate
competitor species identity better fit the data. We conduct an analysis
of the full SCBI plot in Section \ref{results} below.

@@

The resulting RMSE of 0.14 computed using cross-validation is larger
than the earlier RMSE of 0.128, suggesting that models that do not take
the inherent spatial autocorrelation of the data into account generate
error estimates that are overly optimistic; in our case RMSE's that are
too low. We conduct an analysis of the full SCBI plot in Section
\ref{results} below.

\hypertarget{results}{%
\section{Results}\label{results}}

Figure \ref{fig:scbi-simulation}

{\centering \includegraphics[width=1\linewidth]{simulation_results/2021-03-03_scbi_49_shuffles}

\caption{Root mean squared error of models for standard, permuted, and spatial cross-validated error estimates.}\label{fig:scbi-simulation}
\end{figure}

@@

We intend to archive all data and source code for the
\texttt{forestecology} package as well as this manuscript on GitHub at
\url{https://github.com/rudeboybert/forestecology}. This repository will
be versioned and archived on Zenodo upon acceptance. The 2008 and 2014
Smithsonian Conservation Biology Institute census data loaded in Section
\ref{compute-growth} and saved in \texttt{scbi.stem2.csv} and
\texttt{scbi.stem3.csv} are available on GitHub at
\url{https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/tree_main_census/data/census-csv-files}
and have been versioned and archived on Zenodo at
\url{https://doi.org/10.5281/zenodo.2649301}
\citep{gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020}.
